---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, results='hide'}
library (tidyverse)
library(RColorBrewer)
```


#here, where are csv produced by the benchmarking.sh code, as well as the bedtools out produced by the same script. 
```{r, echo=FALSE, warning=FALSE, results='hide'}
path_reference = "/home/tkastylevsky/FASTA_files/repeatmasker/gallus_gallus/reference/galGal6.fa.out.ref.csv"
path_bedtools_out = "/home/tkastylevsky/results/results_bedtools/whole_lib_ensembl_ref.txt"
path_annotation = "/home/tkastylevsky/results/avian_lib_filtered_out/galgal6_ensembl_lib_filtered.fa.out.avian_lib_filtered.csv"
lib_name = "whole_lib_ensembl_filtered"
ref_name = "ref"
```


#création d'une palette afin de garder la meme couleur pour chaque classe de TE. 
bien faire une palette histoire de garder les mêmes couleurs tout du long ! 
```{r, echo=FALSE, warning=FALSE, results='hide'}
my_breaks = c("Simple_repeat", "DNA", "ARTEFACT","LINE", "LTR", "RC", "Low_complexity","rRNA_snRNA_tRNA_scRNA" , "Absent" , "Satellite","SINE",  "Unknown")
myColors <- brewer.pal(12, "Paired")
names(myColors) <- my_breaks
colScale <- scale_fill_manual(name = "class",values = myColors)

```

#loading the data
```{r, echo=FALSE, warning=FALSE, results='hide'}
table_ref = read.table(path_reference, fill = TRUE, header = TRUE)#fill is here because there is a column in the csv that can contain NAs (the column indicating if there is a better TE annotated were the concerned TE is. (I did not remove these TEs))
table_annotation = read.table(path_annotation, fill = TRUE, header = TRUE)
#table_bedtools = read.table(path_bedtools_out)
```


#formating the reference repeatmasker.out
```{r, echo=FALSE, warning=FALSE, results='hide'}
colnames(table_ref) = gsub(ref_name,"B",colnames(table_ref), fixed = TRUE)
table_ref = separate(data = table_ref, col = repeat_class_family, into = c("B_class", "B_family"), sep = "/")

table_ref$B_class = gsub("?",'',table_ref$B_class, fixed = TRUE)#fixed allows to look for the character and not the regular expression
table_ref$B_class = gsub('Unspecified','Unknown',table_ref$B_class)
table_ref$B_class = gsub('rRNA|snRNA|tRNA|scRNA','rRNA_snRNA_tRNA_scRNA',table_ref$B_class)
table_ref$B_class = gsub('Retroposon','LINE',table_ref$B_class)

```

#formating our annotation repeatmasker.out
```{r, echo=FALSE, warning=FALSE, results='hide'}
colnames(table_annotation) = gsub(lib_name,"A",colnames(table_annotation))
table_annotation = separate(data = table_annotation, col = repeat_class_family, into = c("A_class", "A_family"), sep = "/")

table_annotation$A_class = gsub("?",'',table_annotation$A_class, fixed = TRUE)#fixed allows to look for the character
table_annotation$A_class = gsub('Unspecified','Unknown',table_annotation$A_class)
table_annotation$A_class = gsub('rRNA|snRNA|tRNA|scRNA','rRNA_snRNA_tRNA_scRNA',table_annotation$A_class)
table_annotation$A_class = gsub('Retroposon','LINE',table_annotation$A_class)

```

```{r}

table_anno.sum <- table_annotation %>%
    group_by(A_class, TE_avian_lib_filtered_ID) %>%
    tally()


ggplot(data = table_anno.sum, aes(x = n)) +
  geom_histogram() 


table_anno.class.sum  = table_anno.sum %>%
  group_by(A_class) %>%
  tally()



ggplot(data = table_anno.sum, aes(x= A_class, fill = A_class))+
  geom_bar()+
  colScale
ggplot(data = table_annotation,aes(x=A_class,fill=A_class))+
  geom_bar()+
  colScale
```

One code to find them all : making a nice table.
this part is for our annotation, bellow is for the reference.

```{r}
path_OCTFTA="/home/tkastylevsky/results/avian_lib_filtered_out/galgal6_ensembl_avian_filtered_lib_OCTFTA"
library(plyr)
liste_fichiers = list.files(path_OCTFTA)
liste_fichiers = grep("*elem_sorted.csv",liste_fichiers, value = TRUE)

create_table= function(i){
  table_TE = read.table(paste(path_OCTFTA,"/",i,sep=""), header = TRUE, fill = TRUE,comment.char ="")
  
  return(table_TE)
  
}
liste_tables = lapply(liste_fichiers,create_table)
big_table =rbind.fill(liste_tables)

table_insertions = big_table[grep("#",big_table$Score, fixed = TRUE),]# the # rows are the insertions


table_insertions$Family = gsub("DNA/RC","RC",table_insertions$Family)

table_insertions = separate(data = table_insertions, col = Family, into = c("class", "family"), sep = "/")

table_insertions$class = gsub("?",'',table_insertions$class, fixed = TRUE)#fixed allows to look for the character
table_insertions$class = gsub('Unspecified','Unknown',table_insertions$class)
table_insertions$class = gsub('rRNA|snRNA|tRNA|scRNA','rRNA_snRNA_tRNA_scRNA',table_insertions$class)
table_insertions$class = gsub('Retroposon','LINE',table_insertions$class)

ggplot(data = table_insertions, aes(x= class, fill = class))+
  geom_bar()+
  colScale

```

this chunk is to show that some unknowns are labelled as LTR in OCTFTA output.
```{r}
table_LTR_Rmsk = table_annotation[table_annotation$A_class =="LTR",]
table_LTR_OCTFTA = big_table[grep("LTR",big_table$Family),]
table_LTR_OCTFTA = separate(data = table_LTR_OCTFTA, col = Family, into = c("class", "family"), sep = "/")


table_LTR_Rmsk$avian_lib_filtered_ID = NULL
table_LTR_OCTFTA$Num_Assembled = NULL
colnames(table_LTR_Rmsk)=colnames(table_LTR_OCTFTA)


ans = merge(table_LTR_OCTFTA,table_LTR_Rmsk, by = c("Query","Beg.","End."), all.x = TRUE)

table_Unknown_Rmsk = table_annotation[table_annotation$A_class == "Unknown",]
table_Unknown_Rmsk$avian_lib_filtered_ID = NULL
colnames(table_Unknown_Rmsk)=colnames(table_LTR_OCTFTA)

merge_unknown_to_LTR_OCTFTA = merge(table_Unknown_Rmsk[,c('Query', 'Beg.','End.','Element','class','family')],table_LTR_OCTFTA[,c('Query', 'Beg.','End.','Score','class','family')],by = c("Query","Beg.","End."), all.x = TRUE)
```




  
  
```{r}
path_OCTFTA="/home/tkastylevsky/FASTA_files/repeatmasker/gallus_gallus/reference/galgal6_ref_OCTFTA"
library(plyr)
liste_fichiers = list.files(path_OCTFTA)
liste_fichiers = grep("*elem_sorted.csv",liste_fichiers, value = TRUE)

create_table= function(i){
  table_TE = read.table(paste(path_OCTFTA,"/",i,sep=""), header = TRUE, fill = TRUE,comment.char ="")
  
  return(table_TE)
  
}
liste_tables_ref = lapply(liste_fichiers,create_table)
big_table_ref =rbind.fill(liste_tables_ref)

table_insertions_ref = big_table_ref[grep("#",big_table_ref$Score, fixed = TRUE),]# the # rows are the insertions


table_insertions_ref$Family = gsub("DNA/RC","RC",table_insertions_ref$Family)

table_insertions_ref = separate(data = table_insertions_ref, col = Family, into = c("class", "family"), sep = "/")

table_insertions_ref$class = gsub("?",'',table_insertions_ref$class, fixed = TRUE)#fixed allows to look for the character
table_insertions_ref$class = gsub('Unspecified','Unknown',table_insertions_ref$class)
table_insertions_ref$class = gsub('rRNA|snRNA|tRNA|scRNA','rRNA_snRNA_tRNA_scRNA',table_insertions_ref$class)
table_insertions_ref$class = gsub('Retroposon','LINE',table_insertions_ref$class)





```




```{r}
table_insertions_ref$symbol = "ref"
table_insertions$symbol = "anno"

table_sum=rbind(table_insertions,table_insertions_ref)



ggplot(data = table_sum, aes(x= symbol, fill = class))+
  geom_bar(position="dodge")+
  colScale

```
























